---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

```{r}
library(foreign)
library(here)
library(readr)
library(dplyr)
library(tidyverse)
library(readxl)
library(janitor)
library(kableExtra)

xl_data = read_excel('../../inputs/data/GSS/GSS.xlsx')

clean_data <- xl_data |> 
  select(year, id_, age, degree, sex, race) |> 
  filter(degree != ".d:  Do not Know/Cannot Choose") |> 
  filter(degree != ".n:  No answer") |> 
  filter(year > 1997) |>
  filter(year < 2019) 


? filter
  
count_people_degree <- clean_data |> 
  select(year, degree) |>
  group_by(year) |> 
  count(degree)

count_people_degree |>
  ggplot(aes(x = year, y = n)) +
  scale_x_continuous(breaks = seq(1998, 2018, 2)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 65, vjust = 0.6)) +
  geom_point() +
  geom_line() +
  facet_wrap(~degree) +
  labs(
    x = "year",
    y = "n"
  )


degree_by_sex <- clean_data |> 
  select(degree, sex) |>
  group_by(degree, sex) |> 
  count(degree)

degree_by_sex |>
  # select(degree, n , sex) |> 
  ggplot(aes(x = degree, y = n, fill = sex)) +
  scale_y_continuous(breaks = seq(500, 8500, 500)) +
  theme_minimal() +
  geom_col(
    position = "dodge"
  ) +
  labs( #labelling axis
    x = "Degree",
    y = "Number of Degree Holders"
  ) +
  scale_fill_brewer(palette = "BuPu") #setting colour purple colour palette

# degree_by_sex |>
#   ggplot(data=degree_by_sex, aes(x=degree, y=n, fill=sex)) +
#   geom_bar(stat="identity", color="black", position=position_dodge())+
#   theme_minimal()



degree_by_race_filtered <- clean_data |> 
  select(year, degree, race) |>
  filter(degree != "Less than high school") |>
  filter(degree != "High school") |>
  group_by(degree, race) |>
  count(year)

degree_by_race_table <- clean_data |>
  select(year, degree, race) |>
  group_by(year, degree, race) |>
  count(year)

firsttabl <- tabyl(year, degree, race) |> #filtering only needed data
  adorn_totals(c("row", "col")) |> #calculating totals of all selected values
  select("") |> #selecting only years with completed data 
  kable(booktabs = T, caption = "Number of Apprehensions by MHA Apprehension Type from 2014 to 2021",
    col.names = c("MHA Apprehension Types", "2014", "2015", "2016", "2017", "2018", "2019", "2020", "2021"), 
    align = "l"
  )


?summarize

# degree_by_race_associate |
# 
# degree_by_race_bachelors |>   
# 
# 
# degree_by_race_graduate |> 

  
degree_by_race |>
  ggplot(aes(x = year, y = n, color = race)) +
  geom_point() +
  geom_line() +
  theme_minimal() +
  facet_wrap(
    ~degree,
    ncol = 1
  ) +
  scale_x_continuous(breaks = seq(1998, 2018, 2)) +
  scale_y_continuous(breaks = seq(0, 650, 100)) +
  labs(
    x = "Year",
    y = "number of degree holders",
    color = "race"
  ) +
  scale_color_brewer(palette = "Set1")

? ggplot
? facet_wrap




```

```{r}

#| fig-width: 20

```

some more changes

# Introduction

You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point. 



# Data {#sec-data}

Our data is of penguins (@fig-bills).

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (@fig-billssssss). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)

```{r}
#| label: fig-billssssss
#| fig-cap: More bills of penguins
#| echo: false
#| fig-width: 8
#| fig-height: 4

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it. 



# Model

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$ {#eq-bayes}

@eq-bayes seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


